---
title: "BEST Primary Analysis"
subtitle: "ELC Discussion"
author: "John Sperger"
date: "2024-06-12"
always_allow_html: true
format:
  beamer:
    pdf-engine: lualatex
    classoption: professionalfonts, math-style=TeX, spacing = French
    aspectratio: 169
    slide-level: 3
    keep_tex: true
    mainfontoptions:
      - Numbers = Proportional
header-includes:
  - \usepackage{mathtools}
  - \usepackage{graphicx}
  - \usepackage{subfig}
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \usepackage{makecell}
  - \usepackage{wrapfig}
  - \usepackage{rotating}
  - \usepackage[boxed,longend,linesnumbered]{algorithm2e}
  - \usepackage[normalem]{ulem}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{capt-of}
  - \usepackage{hyperref}
  - \usepackage{tikz}
  - \usetikzlibrary{shapes.geometric, arrows}
  - \hypersetup{colorlinks=true, linkcolor=blue, filecolor=magenta, urlcolor=cyan, pdftitle={Overleaf Example}, pdfpagemode=FullScreen}
  - \usepackage[style=verbose,backend=biber,isbn=false,url=false,doi=false,eprint=false,dashed=false]{biblatex}
---

# Preamble
```{r libs_and_scripts}
# Load necessary libraries
library(tidyverse)
library(kableExtra)
library(fabricatr)
library(ggplot2)
library(ggpattern)
library(beanplot)
library(svglite)
library(hrbrthemes)
library(ggpubr)
library(grid)

# Set Default Chunk Options
knitr::opts_chunk$set(echo = FALSE)

# Define constants
seed_val <- 42

colorblind_palette  <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
                         "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
                                        # Load source files
set.seed(seed_val)
source("../code/00_aesthetic_utils.R")
source("../code/01_gen_fake_data.R")
source("../code/02_plot_predictors.R")
source("../code/03_plot_value_comparisons.R")
source("../code/04_plot_dtr_assignments.R")
source("../code/05_plot_forest_subgroups.R")
```


## Overview

The next few slides are a high-level discussion, followed by concrete mock-ups
for each of the two manuscripts.

We will go through the mock tables and figures for the 12- and
24-week manuscripts as they might appear in the manuscripts.

Many of the figures will appear in both manuscripts with different contents.

This is not exhaustive of everything that may be included in appendices.

## Framing

### Language Discussion

This will be the language used in internal ongoing discussions, the design, and the primary analysis manuscripts.

Nonbinding; field and journal specific terms will take precedence for manuscripts.

#### Patient Characteristics

#### Dynamic Treatment Regime (DTR)


### What does a precision medicine analysis look like

### Why both "best patient" and "best treatment" analyses matter
```{r pmed_motivation_plot, fig.height = 3, fig.width = 7}
ThemePlot(CreateExDTRRespDifPlot())
```

The optimal treatment recommendation is "always give meditation" in this
example. Acupuncture gets less effective with age, while meditation gets more effective with age.



Statistical recommendations are always in the context of a set of
options and aren't sufficient to know what predicts treatment response:
\begin{itemize}
\item Features can effect treatment efficacy but not to a sufficient
degree to change the reccommendation

\item A feature may predict treatment response for all treatment options and not
 distinguish who should receive what.

\item Knowing what features predict treatment response isn't enough to know
which treatment to recommend. There is no age (in a human lifespan) where
acupuncture should be recommended in this example despite age reducing the
effectiveness of meditation and increasing the effectiveness of acupuncture.
\end{itemize}


### Scientific framing

# 12-week Paper
## Descriptive summary of trial results
### As-randomized outcomes

## Model Descriptions
### Interpretable Method discussion
Not a table or figure, but an open item

Trees don't seem as popular as expected

### Ensemble model input table
Table of included models
(Appendix material most likely)

### State the optimal DTRs for PEG and PGIC

### Describe the (counterfactual) DTR assignments in the study population
Under the DTR in this population what percentage of people do we expected to be
assigned to each treatment?

## What is the benefit of incorporating patient features?

### Value versus as-randomized outcomes (benefit at the population level)
Figure, expected value under DTR vs. as-randomized

Under this DTR what percentage of people do we expect to augment/maintain/switch?

How does this compare to study assignments and responder categories?

### Stratified by assigned treatment?
Requested by clinicians, but I don’t think this makes sense without subgroups because the populations will be different in the as-randomized and assigned by DTR groups. Note: Might be a way to get at what they want to know, think about it more.

### Stratified by DTR subgroup (and treatment)
As-randomized vs. DTR subgroup outcome comparison

## Predictors of Treatment Response
### Predictor magnitude figure
Figure Variable importance metrics

### Other variable importance metric figure
E.g. Shapley values, one of the importance metrics for trees

## Discussion of results for 12-week PEG and PGIC
### Differences in optimal DTRs?
### Differences in predictors of treatment response?

## Discussion of results for participants with contraindications
### Description of contraindicated populations for each treatment
### DTR assignments by contraindication
### Outcomes for those with contraindications

# 24-week Paper
## Descriptive summary of trial results
### As-randomized outcomes

## Model Descriptions
### Interpretable Method discussion
Listed for completeness, shouldn't require further discussion after the 12-week
##discussion

### Ensemble model input table
Table of included models
(Appendix material most likely)

### State the optimal DTR for 24-week PEG

### Describe the (counterfactual) DTR assignments in the study population
Under the DTR in this population what percentage of people do we expected to be
assigned to each treatment?

## What is the benefit of incorporating patient features?

### Value versus as-randomized outcomes (benefit at the population level)
Figure, expected value under DTR vs. as-randomized

Under this DTR what percentage of people do we expect to augment/maintain/switch?

How does this compare to study assignments and responder categories?

### Stratified by assigned treatment?
Requested by clinicians, but I don’t think this makes sense without subgroups because the populations will be different in the as-randomized and assigned by DTR groups. Note: Might be a way to get at what they want to know, think about it more.

### Stratified by DTR subgroup (and treatment)
As-randomized vs. DTR subgroup outcome comparison

## Predictors of Treatment Response
### Predictor magnitude figure
Figure Variable importance metrics

### Other variable importance metric figure
E.g. Shapley values, one of the importance metrics for trees

## Discussion of results for 12-week and 24-week PEG
### Differences in optimal DTRs?
### Differences in predictors of treatment response?


## Discussion of results for participants with contraindications
### Description of contraindicated populations for each treatment
### DTR assignments by contraindication
### Outcomes for those with contraindications
