---
title: "BEST Primary Analysis"
subtitle: "ELC Discussion"
author: "John Sperger"
date: "2024-06-12"
always_allow_html: true
format:
  beamer:
    pdf-engine: lualatex
    classoption: professionalfonts, math-style=TeX, spacing = French
    aspectratio: 169
    slide-level: 3
    keep_tex: true
    mainfontoptions:
      - Numbers = Proportional
header-includes:
  - \usepackage{mathtools}
  - \usepackage{graphicx}
  - \usepackage{subfig}
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \usepackage{makecell}
  - \usepackage{wrapfig}
  - \usepackage{rotating}
  - \usepackage[boxed,longend,linesnumbered]{algorithm2e}
  - \usepackage[normalem]{ulem}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{capt-of}
  - \usepackage{hyperref}
  - \usepackage{tikz}
  - \usetikzlibrary{shapes.geometric, arrows}
  - \hypersetup{colorlinks=true, linkcolor=blue, filecolor=magenta, urlcolor=cyan, pdftitle={Overleaf Example}, pdfpagemode=FullScreen}
  - \usepackage[style=verbose,backend=biber,isbn=false,url=false,doi=false,eprint=false,dashed=false]{biblatex}
---

# Preamble
## R Header
```{r libs_and_scripts}
# Load necessary libraries
library(tidyverse)
library(kableExtra)
library(fabricatr)
library(ggplot2)
library(ggpattern)
library(beanplot)
library(svglite)
library(hrbrthemes)
library(ggpubr)
library(grid)

# Set Default Chunk Options
knitr::opts_chunk$set(echo = FALSE)

# Define constants
seed_val <- 42

colorblind_palette  <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
                         "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
                                        # Load source files
set.seed(seed_val)
source("../code/00_aesthetic_utils.R")
source("../code/01_gen_fake_data.R")
source("../code/02_plot_predictors.R")
source("../code/03_plot_value_comparisons.R")
source("../code/04_plot_dtr_assignments.R")
source("../code/05_plot_forest_subgroups.R")
```


## Overview

### Overview
The next few slides are a high-level discussion, followed by concrete mock-ups
for each of the two manuscripts.
\vfill

We will go through the mock tables and figures for the 12- and
24-week manuscripts as they might appear in the manuscripts.
\vfill

Many of the figures will appear in both manuscripts with different contents.
\vfill

This is not exhaustive of everything that may be included in appendices.

## Framing Discussion

### Publication Strategy

Co-publish manuscripts for the 12- and 24-week outcomes.

\vfill
There is precedent e.g. the GRADE trial published two results papers in the same issue of the New England Journal of Medicine.
### Language Discussion

This will be the language used in internal ongoing discussions, the design, and the primary analysis manuscripts.
\vfill

Nonbinding; field and journal specific terms will take precedence for manuscripts.
\vfill

#### Patient Features
- Biomarkers
- Phenotypes
- Phenotypic markers
- Patient characteristics
- Patient features


#### Dynamic Treatment Regime (DTR)

### Terms for DTR in the literature
- Individualized Treatment Rule
- Dynamic Treatment Regime
- Treatment policy
- Treatment recommendations
- Policy (in CS and OR)
- Automated clinical decision aid (this is not a 1:1 substitute)
## Fundamental Questions in Precision Medicine
### Fundamental Questions in Precision Medicine
\begin{enumerate}
\item For a specific patient, what is the best treatment?
\vfill

\item For a specific treatment, what patient features are most important for
predicting treatment response?
\end{enumerate}
### Why both "best patient" and "best treatment" analyses matter
```{r pmed_motivation_plot, fig.height = 3, fig.width = 7}
ThemePlot(CreateExDTRRespDifPlot())
```

The optimal treatment recommendation is "always give meditation" in this
example. Acupuncture gets less effective with age, while meditation gets more effective with age.

### Decision-support and Causal Prediction are complementary but distinct

Statistical recommendations are always in the context of a set of
options and aren't sufficient to know what predicts treatment response:
\begin{itemize}
\item Features can effect treatment efficacy but not to a sufficient
degree to change the reccommendation

\item A feature may predict treatment response for all treatment options and not
 distinguish who should receive what.
\end{itemize}

Knowing what features predict treatment response isn't enough to know
which treatment to recommend.

There is no age (in a human lifespan) where acupuncture should be recommended in
this example despite age reducing the effectiveness of meditation and increasing
the effectiveness of acupuncture.

### Analysis as Scientific Questions
\begin{enumerate}
\item "What were the observed outcomes in the trial?"
\vfill

\item "What is the benefit of incorporating patient features?"
\vfill

\item "For a specific patient, what is the best treatment?"
\vfill

\item "For a specific treatment, what patient features predict response to
treatment?"
\vfill

\item "How did each of the above differ for patients with and without contraindications?"

\end{enumerate}

### Analysis as Scientific Questions Details
#### "What were the observed outcomes in the trial?"
Descriptive summary of the as-randomized outcomes
\vfill \pause

#### "What is the benefit of incorporating patient features?"

Hypothesis test of the value of the estimated optimal DTR compared to the
as-randomized outcomes.
\vfill \pause

#### "For a specific patient, what is the best treatment?"

What is the DTR? For an interpretable DTR this can be calculated by hand. More
complex models require a computer.
\vfill

What is the estimated impact of applying the DTR including the proportion of
patients assigned to each treatment and changes in treatment efficacy.
\vfill

If successful, a clinician could reasonably justify applying the DTR published
in the manuscript in their clinic. A further confirmatory trial would be needed
for a higher level of certainty, but these are all treatments in use and if
successful provide evidence supporting the use of the DTR.

### Analysis as Scientific Questions Details
#### "For a specific treatment, what patient features predict response to treatment?"

For each treatment, which features most strongly predict whether a patient will
respond to the treatment? Answered using variable importance plots for
predictors of treatment response.

Important for future research into new treatments and treatment mechanisms as
well as for identifying patients who are unlikely to have successful response to
any of the current treatment options.

Can inform clinical judgment and supplement the recommendation from the DTR.

\vfill \pause


#### "How did each of the above differ for patients with and without contraindications?"

## Future Work
### Future Analyses
  - Additional protocol objectives: patient preferences, deep phenotyping,
    and the 36-week outcome.
  - Investigating "nearly optimal" treatment recommendations and predictors of response to treatment that are nearly as good as those in the primary analysis.
  - Consortium-wide data analyses and integration.

# 12-week Paper
## Descriptive summary of trial results
### As-randomized outcomes

## Model Descriptions
### Interpretable Method discussion
Not a table or figure, but an open item

Trees don't seem as popular as expected

### Ensemble model input table
Table of included models
(Appendix material most likely)

### State the optimal DTRs for PEG and PGIC

### Describe the (counterfactual) DTR assignments in the study population
Under the DTR in this population what percentage of people do we expected to be
assigned to each treatment?

## What is the benefit of incorporating patient features?

### As-randomized Outcome Comparison

Answers the question "What's the benefit of incorporating patient
characteristics into clinical decision making?"

```{r val_comp, fig.height = 3, fig.width = 9, warning = FALSE, message = FALSE}
# Call the function with specified treatment policy names
policy_names <- c("Ensemble", "Following\nRecommended\nTreatments", "Personalize\nStage One Only", "Best Single\nTreatment", "As-randomized")
fake_val_data <- FakeDataForValueComparison(policy_names,
                                            point_estimates = c(.65, .6, .475, .4, .3),
                                            ci_half_widths = c(.15, .15, .15, .15, .1))

fake_val_min_comp <- fake_val_data %>%
    filter(TreatmentPolicy %in% c("Following\nRecommended\nTreatments", "As-randomized"))
val_plot_min <- PlotValueComparison(in.data = fake_val_min_comp) +
    ylab("Treatment Approach")

ThemePlot(val_plot_min)
```

### Open Question: Adjusting for 12-week Comparisons

Protocol primary objctive does not involve the 12-week outcomes

Don't adjust for multiplicity

Don't adjust for multiplicity across the two manuscripts, adjust within the 12-week paper



#### Multiplicity Adjustment Strategies
Step-down test: Only test 12-week if 24-week is significant

### Open Questions (if adjusting for multiplicity)

#### False Discovery Rate (FDR) or Family-Wise Error Rate (FWER) Control

FDR is the proportion of false positives out of all rejected hypotheses

FWER is the probability of making at least one false positive rejection

FWER is stricter and implies FDR control

#### What $\alpha$-level

Standard $\alpha = .05$

### Value versus as-randomized outcomes (benefit at the population level)
Figure, expected value under DTR vs. as-randomized

Under this DTR what percentage of people do we expect to augment/maintain/switch?

How does this compare to study assignments and responder categories?

### Stratified by assigned treatment?
Requested by clinicians, but I don’t think this makes sense without subgroups because the populations will be different in the as-randomized and assigned by DTR groups. Note: Might be a way to get at what they want to know, think about it more.

### Stratified by DTR subgroup (and treatment)
As-randomized vs. DTR subgroup outcome comparison

## Predictors of Treatment Response
### Predictor magnitude figure
Figure Variable importance metrics

### Other variable importance metric figure
E.g. Shapley values, one of the importance metrics for trees

## Discussion of results for 12-week PEG and PGIC
### Differences in optimal DTRs?
### Differences in predictors of treatment response?

## Discussion of results for participants with contraindications
### Description of contraindicated populations for each treatment
### DTR assignments by contraindication
### Outcomes for those with contraindications

# 24-week Paper
## Descriptive summary of trial results
### As-randomized outcomes

## Model Descriptions
### Interpretable Method discussion
Listed for completeness, shouldn't require further discussion after the 12-week
##discussion

### Ensemble model input table
Table of included models
(Appendix material most likely)

### State the optimal DTR for 24-week PEG

### Describe the (counterfactual) DTR assignments in the study population
Under the DTR in this population what percentage of people do we expected to be
assigned to each treatment?

## What is the benefit of incorporating patient features?

### Value versus as-randomized outcomes (benefit at the population level)
Figure, expected value under DTR vs. as-randomized

Under this DTR what percentage of people do we expect to augment/maintain/switch?

How does this compare to study assignments and responder categories?

### Stratified by assigned treatment?
Requested by clinicians, but I don’t think this makes sense without subgroups because the populations will be different in the as-randomized and assigned by DTR groups. Note: Might be a way to get at what they want to know, think about it more.

### Stratified by DTR subgroup (and treatment)
As-randomized vs. DTR subgroup outcome comparison

## Predictors of Treatment Response
### Predictor magnitude figure
Figure Variable importance metrics

### Other variable importance metric figure
E.g. Shapley values, one of the importance metrics for trees

## Discussion of results for 12-week and 24-week PEG
### Differences in optimal DTRs?
### Differences in predictors of treatment response?


## Discussion of results for participants with contraindications
### Description of contraindicated populations for each treatment
### DTR assignments by contraindication
### Outcomes for those with contraindications
