---
title: "R Notebook"
output: html_notebook
---

```{r libs_and_scripts, include = FALSE, echo = FALSE}
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(ggpubr)
library(grid)
# Set Default Chunk Options
knitr::opts_chunk$set(echo = FALSE)

# Define constants
seed_val <- 42

colorblind_palette  <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
            "#F0E442", "#0072B2", "#D55E00", "#CC79A7")



# Load source files
set.seed(seed_val)
source("00_aesthetic_utils.R")
source("01_gen_fake_data.R")
source("02_plot_predictors.R")
source('03_plot_value_comparisons.R')
```

# Preamble

## Discovery Focus Implications

In terms of presentation, what will be in the body of the manuscript versus an
appendix is conditional on the analysis.

# Predictors of Treatment Response
## Single treatments
```{r trt_pred, warning = FALSE, message = FALSE}
pred_summary_df <- FakeDataForPredictorSummary()

#pred_summary_plot <- PlotPredictorSummary(in.data = pred_summary_df)

alt_pred <- PlotPredImportance(in.data = pred_summary_df)

pubr_pred <- PlotPredPubr(in.data = pred_summary_df)

pred_lollipop_plot <- PlotPredictorLollipop(in.data = pred_summary_df)

#ThemePlot(pred_summary_plot)
```
#### All treatment subfigures on the same line, includes predictors for all treatments
```{r pred_pubr, fig.height = 5, fig.width = 8, warning = FALSE, message = FALSE}
pubr_pred
```

#### Grid of treatment subfigures, includes predictors for all treatments
```{r alt_pred, fig.height = 5, fig.width = 8, warning = FALSE, message = FALSE}
ThemePlot(alt_pred)
```

#### Grid of treatment subfigures, only includes predictors relevant for the specific treatment
```{r pred_lollipop_plot, fig.height = 5, fig.width = 8, warning = FALSE, message = FALSE}
ThemePlot(pred_lollipop_plot)
```

## Augmented Interventions
Variable importance plot for predicting response to each of the combination interventions that only includes predictors that aren't included in both of the single treatments.

## Stage Differences
1. Variable importance plot for predicting response to each of the interventions that only includes predictors that are important in stage two but not stage one.

# Interpretable Model Summary
## Purely Data-driven

## Structured GAM
Identify a small number of predictor categories.

Analagous to multi-scale diagnostic tools (wrong phrasing, look up)

# Comparing Treatment Policies
## Minimize number of comparisons
```{r val_comp, fig.height = 4, fig.width = 6, warning = FALSE, message = FALSE}
# Call the function with specified treatment policy names
policy_names <- c("Ensemble", "Interpretable", "Personalize\nStage One Only", "Best Single\nTreatment", "As-randomized")
fake_val_data <- FakeDataForValueComparison(policy_names,
                                            point_estimates = c(.65, .6, .475, .4, .3),
                                            ci_half_widths = c(.15, .15, .15, .15, .1))

fake_val_min_comp <- fake_val_data %>% filter(TreatmentPolicy %in% c("Ensemble", "Interpretable","As-randomized"))
val_plot_multi <- PlotValueComparison(in.data = fake_val_data)
val_plot_min <- PlotValueComparison(in.data = fake_val_min_comp)

ThemePlot(val_plot_min)
```
## Include additional comparisons
```{r val_comp_multi, fig.height = 4, fig.width = 6, warning = FALSE, message = FALSE}
ThemePlot(val_plot_multi)
```
# Characterizing the Treatment Subgroups

