---
title: "R Notebook"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  ioslides_presentation:
    toc: true
    toc_depth: 2
    slide_level: 3
    widescreen: true
    smaller: true
  html_document:
    toc: true
    toc_depth: 2
    theme: united
  pdf_document:
    keep_tex: true
    toc: true
    toc_depth: 2
  beamer_presentation:
    toc: true
    toc_depth: 2
    slide_level: 3
---

```{r libs_and_scripts, include = FALSE, echo = FALSE}
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(ggpubr)
library(grid)
library(networkD3)
# Set Default Chunk Options
knitr::opts_chunk$set(echo = FALSE)

# Define constants
seed_val <- 42

colorblind_palette  <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
            "#F0E442", "#0072B2", "#D55E00", "#CC79A7")



# Load source files
set.seed(seed_val)
source("00_aesthetic_utils.R")
source("01_gen_fake_data.R")
source("02_plot_predictors.R")
source('03_plot_value_comparisons.R')
source('04_plot_dtr_assignments.R')
source('05_plot_forest_subgroups.R')
```

# Predictors of Treatment Response
### Single treatments
```{r trt_pred, warning = FALSE, message = FALSE}
pred_summary_df <- FakeDataForPredictorSummary()

#pred_summary_plot <- PlotPredictorSummary(in.data = pred_summary_df)

alt_pred <- PlotPredImportance(in.data = pred_summary_df)

pubr_pred <- PlotPredPubr(in.data = pred_summary_df)

pred_lollipop_plot <- PlotPredictorLollipop(in.data = pred_summary_df)

#ThemePlot(pred_summary_plot)
```
#### All treatment subfigures on the same line, includes predictors for all treatments
```{r pred_pubr, fig.height = 5, fig.width = 8, warning = FALSE, message = FALSE}
pubr_pred
```

#### Grid of treatment subfigures, includes predictors for all treatments
```{r alt_pred, fig.height = 5, fig.width = 8, warning = FALSE, message = FALSE}
ThemePlot(alt_pred)
```

#### Grid of treatment subfigures, only includes predictors relevant for the specific treatment
```{r pred_lollipop_plot, fig.height = 5, fig.width = 8, warning = FALSE, message = FALSE}
ThemePlot(pred_lollipop_plot)
```

### Augmented Interventions
Variable importance plot for predicting response to each of the combination interventions that only includes predictors that aren't included in both of the single treatments.

### Stage Differences
1. Variable importance plot for predicting response to each of the interventions that only includes predictors that are important in stage two but not stage one.
2. If there are very few or no differences, a combined variable importance plot for stages one and two with some designation of whether the predictor is for stage one, stage two, or both.

```{r temp, warning = FALSE, message = FALSE, include = FALSE}

   if(FALSE){"
## Purely Data-driven

## Structured GAM
Identify a small number of predictor categories.

Analagous to multi-scale diagnostic tools (wrong phrasing, look up)"
}
```
## Estimated Treatment Policies
Mocking up in `tikz`

1. Ensemble (approximated)
2. Interpretable
3. Comparison of assignments by stage and treatment policy

### Proportion of Assigned Treatments - Trial Design Rule
```{r dtr_assignments, fig.height = 4, fig.width = 6, warning = FALSE, message = FALSE}
#
dtr_assignments <- FakeDataForDTRAssignments()

PlotDTRAssignments(dtr_assignments)
```


### Proportion of Assigned Treatments - Complex ACT rule
```{r dtr_assignments_complex, fig.height = 4, fig.width = 6, warning = FALSE, message = FALSE}
#
dtr_assignments <- FakeDataForDTRAssignments(stage1_trt_probs = c(0.6, 0.1, 0.2, 0.1),
                                             act_action_probs = c(0.6, 0.1, 0.3),
                                             dulox_action_probs = c(0.1, 0.8, 0.1),
                                             ebem_action_probs = c(0.1, 0.4, 0.5),
                                             esc_action_probs = c(0.2, 0.8, 0))

PlotDTRAssignments(dtr_assignments)
```


### Proportion of Assigned Treatments - Simple Rule (for one subgroup - perhaps)
```{r dtr_assignments_simple, fig.height = 4, fig.width = 6, warning = FALSE, message = FALSE}
#
dtr_assignments <- FakeDataForDTRAssignments(stage1_trt_probs = c(0.6, 0, 0, 0.4),
                                             act_action_probs = c(0.6, 0.4, 0),
                                             dulox_action_probs = c(0.1, 0.8, 0.1),
                                             ebem_action_probs = c(0.1, 0.4, 0.5),
                                             esc_action_probs = c(0.5, 0.5, 0))

PlotDTRAssignments(dtr_assignments)
```


## Comparing Treatment Policies
### Small number of comparisons
```{r val_comp, fig.height = 4, fig.width = 6, warning = FALSE, message = FALSE}
# Call the function with specified treatment policy names
policy_names <- c("Ensemble", "Interpretable", "Personalize\nStage One Only", "Best Single\nTreatment", "As-randomized")
fake_val_data <- FakeDataForValueComparison(policy_names,
                                            point_estimates = c(.65, .6, .475, .4, .3),
                                            ci_half_widths = c(.15, .15, .15, .15, .1))

fake_val_min_comp <- fake_val_data %>% filter(TreatmentPolicy %in% c("Ensemble", "Interpretable","As-randomized"))
val_plot_multi <- PlotValueComparison(in.data = fake_val_data)
val_plot_min <- PlotValueComparison(in.data = fake_val_min_comp)

ThemePlot(val_plot_min)
```
### Include additional policies
```{r val_comp_multi, fig.height = 4, fig.width = 6, warning = FALSE, message = FALSE}
ThemePlot(val_plot_multi)
```

## Characterizing the Treatment Subgroups

### Forest Plot for Subgroups

```{r}
dtr_data <- tibble(
  subgroup = c("Overall", "Age <=65 yr", "Age > 65", "Years cLBP >5", "Years cLBP <=5",
               "Anxiety +", "Anxiety -"), # ... continue for all subgroups
  value = c(1, 1.1, 0.6, 0.7, 1.2, 1.3, 0.65),
  method = 'Personalized Treatments'
) %>%
  mutate(subgroup = factor(subgroup,
                           levels = rev(c("Overall", subgroup[which(subgroup != 'Overall')]))))

trial_average_data <- tibble(
  subgroup = c("Overall", "Age <=65 yr", "Age > 65", "Years cLBP >5", "Years cLBP <=5",
               "Anxiety +", "Anxiety -"), # ... continue for all subgroups
  value = c(0.85, 0.8, 0.9, 1, 0.72, 1, 0.8),
  method = 'Trial Average'
) %>%
  mutate(subgroup = factor(subgroup,
                           levels = rev(c("Overall", subgroup[which(subgroup != 'Overall')]))))


pdat <- bind_rows(dtr_data, trial_average_data)

segments_data <- pdat %>%
    pivot_wider(id_cols = subgroup,names_from = method,
                values_from = value) %>%
    janitor::clean_names() %>%
    mutate(
      diff = personalized_treatments - trial_average,
      color = ifelse(diff > 0, "Positive", "Negative")
    )

PlotSubgroupForest(value_data = pdat, segments_data = segments_data)

```


